---
title: "problem 2"
author: "Richard Marks"
date: "2/10/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(mosaic)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
load("~/.Rdata")
plot(creatclear ~ age, data = creatinine)
```
```{r}
lml = lm(creatclear ~ age, data = creatinine)
plotModel(lml)
coef(lml)
```



2. How does creatinine clearance rate change with age?
Every year someone ages their creatnine clearance rate decreases by .62 mL/(min per year)



```{r}
data55 = data.frame(age= c(55))
predict(lml,data55)

```



What creatinine clearance rate should we expect, on average, for a 55-year-old?
113.72 mL/(min per year)




```{r}
data40 = data.frame(age= c(40))
predict(lml,data40)
data60 = data.frame(age= c(60))
predict(lml,data60)
```



3.Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112? A 40 year old with a rate of 135 (12mL/(min per year above the predicted level) is healthier than a 60 year old with a rate of 112 (1.4mL/(min per year above the predicted level)